Adaptive Approximation for Multivariate Linear Problems \\ with Inputs Lying in a Cone

### 2018-11-02 RICAM2018NovTalk.tex
Updating changes
adding

\begin{frame}{Multivariate Linear Problems}
\vspace{-3ex}
\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
Given $f \in \cf$ find $S(f) \in \cg$, where \\
$\sol: \cf \to \cg$ is linear, e.g., 
\begin{gather*}
    S(f) = \int_{\reals^d} f(\vx) \, \varrho(\vx) \, \dif \vx\\
    S(f) = f\\
    S(f) = \frac{\partial f}{\partial x_1} \\
    - \nabla^2 S(f) = f, \ \  S(f) = 0 \text{ on boundary}
\end{gather*}
&
\vspace{-5ex}
\alert{Successful algorithms}
\vspace{-3ex}
\begin{multline*}
    \ca(\cc) : = \{\app: \cc \times (0,\infty) \to \cg \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \ \ \forall f \in \cc \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

@@

\vspace{-4ex}
\begin{gather*}
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n f(\vx_i) g_i, \quad g_i \in \cg \\
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n \hf_i g_i, \quad g_i \in \cg \\
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n L_i(f) g_i, \quad g_i \in \cg \\
\end{gather*}

\vspace{-5ex}
\hfill \hfill \alert{$n = ?$, $\cc$ is not a ball}

@@

\begin{frame}{Cones}
\vspace{-4ex}
\begin{tabular}{>{\centering}m{0.4\textwidth}@{\qquad}>{\centering}m{0.4\textwidth}}
     \largescoop \hspace{-3cm}\raisebox{-4ex}{\color{red}\fontsize{100}{120}\selectfont $\times$} & 
      \largecone \tabularnewline
      Ball $\cb_{\rho} := \{f \in \cf : \norm[\cf]{f} \le \rho \}$ &
      \hspace{2cm} (Non-Convex) Cone $\cc$
\end{tabular}

@@

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_4.png} 
	\tabularnewline[-7ex]
	Legendre
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_4.png} 
\tabularnewline[-7ex]
Chebyshev \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_4.png} 
	\tabularnewline[-7ex]
	Legendre
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_4.png} 
\tabularnewline[-7ex]
Chebyshev \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}{Series Spaces}

@@

\begin{align*}
    \cf &:= \left \{ f = \sum_{\vi \in \cn} \hf(\vi) u_{\vi} : \norm[\cf]{f} := \bignorm[r,\vlambda]{\hf} : = \norm[r]{\left(\frac{\bigabs{\hf(\vi)}}{\lambda_{\vi}} \right)_{\vi \in \cn}} \right \} \\
    \cg &: = \left \{ g = \sum_{\vi \in \cn} \hg(\vi) v_{\vi} : \norm[\cg]{g} := \bignorm[t]{\hg}\right \}, \qquad S(u_{\vi}) = v_{\vi}, \\
    \text{E.g., } & \text{ Function recovery: } S(u_{\vi}) = u_{\vi} = v_{\vi} \\
    & \text{ Integration: } S(u_{\vi}) = \int_{[0,1]^d} u_\vi (\vx) \, \dif \vx  = v_\vi\\
    \norm[\cg]{S(f) - A(f,\varepsilon)} & \le \sum_{\vi \in \cn} \bigabs{\hf(\vi)} \norm[\cg]{v_\vi - \Sapp(u_{\vi},n)}
\end{align*}
If $\gamma_{\vi_1} \ge \gamma_{\vi_2} \ge \cdots$, then $S_{\textup{app},n}(f) = \sum_{j=1}^n \hf(\vi_j) v_{\vi_j}$ is an optimal building block, and 
\begin{equation*}
    \norm[t]{S(f) - S_{\textup{app},n}(f)} = \Bignorm[r']{\left(\hf(\vi_j)\right)_{j=n+1}^{\infty}} \le \bignorm[r,\vgamma]{\hf} \Bignorm[qq'/(q-q')]{\left(\gamma_{\vi_j}\right)_{j=n+1}^{\infty}}
\end{equation*}

@@

>>>>

\begin{frame}{Multivariate Linear Problems}
\vspace{-3ex}
\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
Given $f \in \cf$ find $S(f) \in \cg$, where \\
$\sol: \cf \to \cg$ is linear, e.g., 
\begin{gather*}
    S(f) = \int_{\reals^d} f(\vx) \, \varrho(\vx) \, \dif \vx\\
    S(f) = f\\
    S(f) = \frac{\partial f}{\partial x_1} \\
    - \nabla^2 S(f) = f, \ \  S(f) = 0 \text{ on boundary}
\end{gather*}
&
\vspace{-7ex}
\alert{Successful algorithms}
\vspace{-3ex}
\begin{multline*}
    \ca(\cc) : = \{\app: \cc \times (0,\infty) \to \cg \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \ \ \forall f \in \cc \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

@@

\vspace{-4ex}
\begin{gather*}
    \Sapp(f,n) = \sum_{i=1}^n f(\vx_i) g_i, \quad g_i \in \cg \\
    \Sapp(f,n) = \sum_{i=1}^n \hf_i g_i, \quad g_i \in \cg \\
    \Sapp(f,n) = \sum_{i=1}^n L_i(f) g_i, \quad g_i \in \cg \\
\end{gather*}

\vspace{-5ex}
\hfill \hfill \alert{$\app(f,\varepsilon) = \Sapp(f,n) + $ stopping criterion}\newline
\phantom{a} \hfill \hfill \alert{$\cc$ is not a ball}

@@

\begin{frame}{Cones}
\vspace{-4ex}
\begin{tabular}{>{\centering}m{0.4\textwidth}@{\qquad}>{\centering}m{0.4\textwidth}}
     \largescoop \hspace{-3cm}\raisebox{-4ex}{\color{red}\fontsize{100}{120}\selectfont $\times$} & 
      \largecone \tabularnewline
      Ball $\cb_{\rho} := \{f \in \cf : \norm[\cf]{f} \le \rho \}$ &
      \hspace{2cm} (Non-Convex) Cone $\cc$
\end{tabular}
answer those who like balls

@@

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_4.png} 
	\tabularnewline[-7ex]
	Cosine \& Sine
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/CosineSine_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_4.png} 
\tabularnewline[-7ex]
Walsh \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Walsh_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_4.png} 
	\tabularnewline[-7ex]
	Legendre
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Legendre_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_4.png} 
\tabularnewline[-7ex]
Chebyshev \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{ProgramsImages/Chebyshev_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}{Integration Using Function Values \footfullcite{HicJim16a,JimHic16a}}

@@

\begin{align*}
    \cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \bignorm[1,\vlambda]{\hf} : = \norm[1]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \}\\
    u_{\vj} &= \text{Walsh or Cosine/Sine} \qquad S(u_{\vj}) = \int_{[0,1]^d} u_\vj (\vx) \, \dif \vx = \delta_{\vj,\vzero}, \qquad \cg : = \reals \\
    \Sapp(f,n) &= \frac 1n \sum_{i=1}^n f(\vx_i) = \sum_{\vj \in \text{dual set}} \hf(\vj), \qquad \{\vx_i\}_{i=1}^n = \text{net or lattice} \quad   \\
    \abs{S(f) - A(f,\varepsilon)} & \le \sum_{\vj \in \text{dual set}} \bigabs{\hf(\vj)}
\end{align*}

@@

\begin{frame}{But I Like Balls \smallscoop!}

\vspace{-5ex}
How might you construct an adaptive algorithm if you insist on using \alert{balls}?

\vspace{-2ex}
\begin{description}
    \item[Step 1] Start with a default radius $\rho$, and assume input $f \in \cb_{\rho}$

    \item[Step 2] Choose $n$ large enough so that 
    \[
    \norm[\cg]{S(f) - \Sapp(f,n)} \le \varepsilon \qquad \forall f \in \cb_{\rho}, \qquad \text{where } \Sapp(f,n) = \sum_{i=1}^n L_i(f) g_i
    \]

    \item[Step 3] Let $\tf \in \cf$ be the \alert{minimum} norm interpolant of the data $L_1(f), \ldots, L_n(f)$.

    \item[Step 4]  If $C\bignorm[\cf]{\tf} \le \rho$ for some preset inflation factor, $C$, then return $A(f,\varepsilon) = \Sapp(f,n)$.  Otherwise, choose $\rho = 2C\bignorm[\cf]{\tf}$, and go to Step 2.
\end{description}

This algorithm succeeds for the \alert{cone} \smallcone defined as those functions in $\cf$ whose norms are not much larger than their minimum norm interpolants.
